### 2017-07-07 motivation.tex
edits

\begin{figure}[t]
  \centering
	\begin{subfigure}{0.4\textwidth}
		\includegraphics[width=0.9\textwidth]{Figures/schema1}
    \caption{A valid TPC-C database. The only existing order belongs
      to the district with \C{d\_id}=11. Its id (\C{o\_id}) is one
      less than the district's \C{d\_next\_o\_id}, and its order
      count (\C{o\_ol\_cnt}) is equal to the number of order line
      records whose \C{ol\_o\_id} is equal to the order's id.  }
		\label{fig:tpcc_db1}
	\end{subfigure}
	\begin{subfigure}{0.52\textwidth}
		\includegraphics[width=0.9\textwidth]{Figures/schema2}
    \caption{The database in Fig.~\ref{fig:tpcc_db1} after correctly
      executing a \C{new\_order} transaction. A new order record is
      added whose \C{o\_id} is equal to the \C{d\_next\_o\_id} from
      Fig.~\ref{fig:tpcc_db1}. The district's \C{d\_next\_o\_id} is
      incremented. The order's \C{o\_ol\_cnt} is 2, reflecting the
      actual number of order line records whose \C{ol\_o\_id} is equal
      to the order's id (2).}
		\label{fig:tpcc_db2}
	\end{subfigure}

\caption{Database schema of TPC-C's order management system.
  The naming
  convention indicates primary keys and foreign keys. For e.g.,
  \C{ol\_id} is the primary key column of the order line table,
  whereas \C{ol\_o\_id} is a foreign key that refers to the \C{o\_id}
  column of the order table.}
\label{fig:schema}
\end{figure}

@@

Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundess of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes.
Under weak isolation\footnote{Weak isolation does not violate
  atomicity as long as the witnessed effects are those of committed
  transactions}, however, interferences of various kinds are
permitted, leading to executions superficially similar to executions
permitted by concurrent (racy) programs~\cite{GHE15,HPQ+15}.
\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits. Nonetheless, reading the same value for
  \C{d\_next\_o\_id} prompts both instances to add \C{Order} records
  with same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\end{wrapfigure}
To illustrate, consider the behavior of the \C{new\_order} transaction
when executed with a \emph{Read Committed} (RC) isolation level, the
default isolation level in 8 of the 18 databases studied
in~\ref{bailishat}.  An RC transaction is isolated from \emph{dirty
  writes}, i.e., writes of uncommitted transactions, but is allowed to
witness the writes of concurrent transactions as soon as they are
committed. Thus, with two concurrent instances of the \C{new\_order}
transaction (call them $T_1$ and $T_2$), both concurrently placing new
orders for different customers under the same district (\C{d\_id}), RC
isolation allows the execution shown in
Fig.~\ref{fig:new_order_execs}.

@@

In this paper, we adopt a different approach that \emph{lifts}
isolation semantics (\emph{not} their implementations) to the
application layer, providing a principled framework to simultaneously
reason about application invariants and isolation properties.  To
illustrate this idea informally, consider how we might verify that
\C{new\_order} is sound when executed under \emph{Snapshot Isolation}
(SI), a stronger isolation level than RC. Snapshot isolation allows
transactions to be executed against a private snapshot of the
database, thus admitting concurrency, but it also requires that there
not be any write-write conflicts among concurrent transactions when
they commit. Write-write conflicts can be eliminated in various ways,
e.g., through conflict detection followed by a rollback, or through
exclusive locks, or a combination of both.  For instance, one possible
implementation of SI, close to the one used by
PostgreSQL~\cite{postgres-ssi}, executes a transaction against its
private snapshot of the database, but obtains exclusive locks on the
actual records in the database before performing writes. A write is
performed only if the record that is to be written has not already
been updated by a concurrent transaction. Conflicts are resolved by
abort and roll back.

@@

Implementations of SI on real databases (e.g., PostgreSQL) are
complicated, often running into thousands of lines of code.
Nonetheless, the semantics of SI, in terms of how it effects
transitions on the database state, can be captured in a fairly simple
model. First, effects induced by one transaction are not visible to
another concurrently executing one (call it $T$), during $T$'s
execution. Thus, from $T$'s perspective, the global state does not
change during its execution.  More formally, for every operation
performed by $T$, the global state $T$ witness before ($\stg$) and
after ($\stg')$ executing the operation is the same ($\stg'=\stg$).
After $T$ finishes execution, it commits its changes to the actual
database, which may have already incorporated the effects of concurrent
transactions. In executions where $T$ successfully commits, concurrent
transactions are guaranteed to not be in write-write conflict with
$T$. Thus, if $\stg$ is the global state that $T$ witnessed when it
finished execution (the snapshot state), and
$\stg'$ is the state to which $T$ commits, then the difference between
$\stg$ and $\stg'$ should not result in a write-write conflict with $T$. To
concretize this notion, let the database state be a map from
transaction variables to values, and let $\stl$ denote a
transaction-local log that maps the variables being written to their
updated values.  The absence of write-write conflicts between $T$
and the diff between $\stg$ and $\stg'$ can be expressed as: $\forall
x\in\mathit{dom}(\delta)$, $\stg'(x) = \stg(x)$.  In other words, the
semantics of SI can be captured as an axiomatization over transitions
of the database state ($\Delta \longrightarrow \Delta'$) during the
lifetime an SI transaction ($T$):
\begin{itemize}
  \item While $T$ executes, $\Delta' = \Delta$.
  \item After $T$ finishes execution, but before it commits its local
    state $\delta$, $\forall(x\in\delta).~\Delta'(x) = \Delta(x)$.
\end{itemize}

@@

Fig.~\ref{fig:foreach_code} shows a (simplified) snippet of code taken
from Fig.~\ref{fig:new_order_code}. Some irrelevant expressions have
been replaced with constants (\C{k1} to \C{k3}).  The body of the loop
executes a SQL update followed by an insert.  Recall that a transaction
reads from the global database ($\Delta$), and writes to a
transaction-local database ($\delta$) before committing these updates. An update
statement filters the tuples that match the search criteria from $\Delta$
and computes the updated tuples that are to be
added to the local database. Thus, its state transformer (call it
$\T_U$) is the following function on sets\footnote{$\bind$ has higher
precedence than $\cup$.}:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \stg \bind(\lambda s. 
      \itel{{\sf table}(s) = \C{Stock} \conj 
                s.\C{s\_i\_id} = \C{item\_req.ol\_i\_id}\\\hspace*{1.3in}} 
           {\{\{s \with \C{s\_qty} = \C{k1}\}\}\\\hspace*{1.3in}}
           {\emptyset}
\end{array}
\end{smathpar}
% \begin{smathpar}
% \begin{array}{c}
%   \lambda\Delta.\lambda\delta.\, \delta \;\cup\; 
%       \Delta \,\bind\, (\lambda s.\, 
%           \C{if}\;{\C{stock}(s) \conj 
%                    \C{s.s\_i\_id} = \C{item\_req.ol\_i\_id}}\\
%            \hspace*{0.9in}
%            \C{then}\;{\{\{s \;\C{with}\; \C{s\_qty} = \C{k1}\}\}}\;
%            \C{else}\;{\emptyset})
% \end{array}
% \end{smathpar}
% \begin{verbatim}
%   Rmem(σ'(s')) = Rmem(σ(s')) U 
%                  Rmem[\x. Rmem[\y. if y<x then {y+x} else {}](σ(s)) U {x}](xs)
% \end{verbatim}
% Observe that
% $\T_U(\Delta,\delta)$ is of the form $\delta \,\cup\,
% F_U(\Delta)$. 
The transformer ($\T_I(\Delta,\delta)$) for the subsequent \C{insert}
statement can be similarly constructed:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \{\C{ol\_o\_id}=\C{k2};\, 
        \C{ol\_d\_id}=\C{k3};\, \C{ol\_i\_id}=\C{item\_req.ol\_i\_id};\, 
        \C{ol\_qty}=\C{item\_req.ol\_qty}\}
\end{array}
\end{smathpar}
Observe that both transformers are of the form $T(\Delta,\delta) =
\stl \cup F(\stg)$, where $F$ is a function that returns the set of
records added by the transformer to the transaction-local database
($\stl$). Let $\F_U$ and $\F_I$ be the corresponding functions for $\T_U$
and $\T_I$ shown above. The state transformed of the loop body in
Fig.~\ref{fig:new} can be expressed as the following composition of
$\F_U$ and $\F_I$:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \F_U(\stg) \cup \F_I(\stg)
\end{array}
\end{smathpar}
The transformer for the loop itself can now be computed to be:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \C{item\_reqs}\bind
      (\lambda\C{item\_req}.~ \F_U(\stg) \cup \F_I(\stg))
\end{array}
\end{smathpar}
Observe that the structure of the transformer mirrors the structure of
the program itself. In particular, SQL statements become set
operations, and the \C{DB} monad's $\bind$ becoms set monadic $\bind$.
Thus, state transformer inference is completely mechanical. The
advantage of inferring such transformers is that we can now make use
of a semantics-preserving translation from the domain of sets equipped
with $\bind$ to first-order logic, allowing us to leverage SMT solvers
for automatic proofs without having to infer loop invariants, or
provide intermediate assertions.  Sec.~\ref{sec:inference} describes
this translation. In the exposition thus far, we assumed $\Delta$
remains invaraint, which is clearly not the case when we admit
concurrency.  Necessary concurrency extensions of the state
transformer semantics to deal with interference is also covered in
Sec.~\ref{sec:inference}.  The following two sections, however, first
focus on laying the theoretical foundations for reasoning about weak
isolation.

>>>>

\begin{figure}[t]
  \centering
	\begin{subfigure}{0.4\textwidth}
	  \includegraphics[width=0.9\textwidth]{Figures/schema1}
    \caption{A valid TPC-C database. The only existing order belongs
      to the district with \C{d\_id}=11. Its id (\C{o\_id}) is one
      less than the district's \C{d\_next\_o\_id}, and its order
      count (\C{o\_ol\_cnt}) is equal to the number of order line
      records whose \C{ol\_o\_id} is equal to the order's id.  }
		\label{fig:tpcc_db1}
	\end{subfigure}
  \hspace*{.4in}
	\begin{subfigure}{0.52\textwidth}
		\includegraphics[width=0.9\textwidth]{Figures/schema2}
    \caption{The database in Fig.~\ref{fig:tpcc_db1} after correctly
      executing a \C{new\_order} transaction. A new order record is
      added whose \C{o\_id} is equal to the \C{d\_next\_o\_id} from
      Fig.~\ref{fig:tpcc_db1}. The district's \C{d\_next\_o\_id} is
      incremented. The order's \C{o\_ol\_cnt} is 2, reflecting the
      actual number of order line records whose \C{ol\_o\_id} is equal
      to the order's id (2).}
		\label{fig:tpcc_db2}
   	\end{subfigure}
\caption{Database schema of TPC-C's order management system.
  The naming
  convention indicates primary keys and foreign keys. For e.g.,
  \C{ol\_id} is the primary key column of the order line table,
  whereas \C{ol\_o\_id} is a foreign key that refers to the \C{o\_id}
  column of the order table.}
\label{fig:schema}
\end{figure}

@@

Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundess of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes.
Under weak isolation\footnote{Weak isolation does not violate
  atomicity as long as the witnessed effects are those of committed
  transactions}, however, interferences of various kinds are
permitted, leading to executions superficially similar to executions
permitted by concurrent (racy) programs~\cite{GHE15,HPQ+15}.
To illustrate, consider the behavior of the \C{new\_order} transaction
\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits. Nonetheless, reading the same value for
  \C{d\_next\_o\_id} prompts both instances to add \C{Order} records
  with same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\end{wrapfigure}
when executed with a \emph{Read Committed} (RC) isolation level, the
default isolation level in 8 of the 18 databases studied
in~\cite{bailishotos}.  An RC transaction is isolated from \emph{dirty
  writes}, i.e., writes of uncommitted transactions, but is allowed to
witness the writes of concurrent transactions as soon as they are
committed. Thus, with two concurrent instances of the \C{new\_order}
transaction (call them $T_1$ and $T_2$), both concurrently placing new
orders for different customers under the same district (\C{d\_id}), RC
isolation allows the execution shown in Fig.~\ref{fig:new_order_exec}.

@@

In this paper, we adopt a different approach that \emph{lifts}
isolation semantics (\emph{not} their implementations) to the
application layer, providing a principled framework to simultaneously
reason about application invariants and isolation properties.  To
illustrate this idea informally, consider how we might verify that
\C{new\_order} is sound when executed under \emph{Snapshot Isolation}
(SI), a stronger isolation level than RC. Snapshot isolation allows
transactions to be executed against a private snapshot of the
database, thus admitting concurrency, but it also requires that there
not be any write-write conflicts among concurrent transactions when
they commit. Write-write conflicts can be eliminated in various ways,
e.g., through conflict detection followed by a rollback, or through
exclusive locks, or a combination of both.  For instance, one possible
implementation of SI, close to the one used by
PostgreSQL~\cite{postgresiso}, executes a transaction against its
private snapshot of the database, but obtains exclusive locks on the
actual records in the database before performing writes. A write is
performed only if the record that is to be written has not already
been updated by a concurrent transaction. Conflicts are resolved by
abort and roll back.

@@

As this discussion indicates, implementations of SI on real databases
such as PostgresSQL are highly complicated, often running into
thousands of lines of code.  Nonetheless, the semantics of SI, in
terms of how it effects transitions on the database state, can be
captured in a fairly simple model. First, effects induced by one
transaction are not visible to another concurrently executing one
(call it $T$), during $T$'s execution. Thus, from $T$'s perspective,
the global state does not change during its execution.  More formally,
for every operation performed by $T$, the global state $T$ witness
before ($\stg$) and after ($\stg')$ executing the operation is the
same ($\stg'=\stg$).  After $T$ finishes execution, it commits its
changes to the actual database, which may have already incorporated
the effects of concurrent transactions. In executions where $T$
successfully commits, concurrent transactions are guaranteed to not be
in write-write conflict with $T$. Thus, if $\stg$ is the global state
that $T$ witnessed when it finished execution (the snapshot state),
and $\stg'$ is the state to which $T$ commits, then the difference
between $\stg$ and $\stg'$ should not result in a write-write conflict
with $T$. To concretize this notion, let the database state be a map
from transaction variables to values, and let $\stl$ denote a
transaction-local log that maps the variables being written to their
updated values.  The absence of write-write conflicts between $T$ and
the diff between $\stg$ and $\stg'$ can be expressed as: $\forall
x\in\mathit{dom}(\delta)$, $\stg'(x) = \stg(x)$.  In other words, the
semantics of SI can be captured as an axiomatization over transitions
of the database state ($\Delta \longrightarrow \Delta'$) during the
lifetime an SI transaction ($T$):
\begin{itemize}
  \item While $T$ executes, $\Delta' = \Delta$.
  \item After $T$ finishes execution, but before it commits its local
    state $\delta$, $\forall(x\in\delta).~\Delta'(x) = \Delta(x)$.
\end{itemize}

@@

Fig.~\ref{fig:foreach_code} shows a (simplified) snippet of code taken
from Fig.~\ref{fig:new_order_code}. Some irrelevant expressions have
been replaced with constants (\C{k1} to \C{k3}).  The body of the loop
executes a SQL update followed by an insert.  Recall that a transaction
reads from the global database ($\Delta$), and writes to a
transaction-local database ($\delta$) before committing these updates. An update
statement filters the tuples that match the search criteria from $\Delta$
and computes the updated tuples that are to be
added to the local database. Thus, its state transformer (call it
$\T_U$) is the following function on sets\footnote{$\bind$ has higher
precedence than $\cup$.}:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \stg \bind(\lambda s. 
      \itel{{\sf table}(s) = \C{Stock} \conj 
                s.\C{s\_i\_id} = \C{item\_req.ol\_i\_id}\\\hspace*{1.3in}} 
           {\{\{s \with \C{s\_qty} = \C{k1}\}\}\\\hspace*{1.3in}}
           {\emptyset}
\end{array}
\end{smathpar}
% \begin{smathpar}
% \begin{array}{c}
%   \lambda\Delta.\lambda\delta.\, \delta \;\cup\; 
%       \Delta \,\bind\, (\lambda s.\, 
%           \C{if}\;{\C{stock}(s) \conj 
%                    \C{s.s\_i\_id} = \C{item\_req.ol\_i\_id}}\\
%            \hspace*{0.9in}
%            \C{then}\;{\{\{s \;\C{with}\; \C{s\_qty} = \C{k1}\}\}}\;
%            \C{else}\;{\emptyset})
% \end{array}
% \end{smathpar}
% \begin{verbatim}
%   Rmem(σ'(s')) = Rmem(σ(s')) U 
%                  Rmem[\x. Rmem[\y. if y<x then {y+x} else {}](σ(s)) U {x}](xs)
% \end{verbatim}
% Observe that
% $\T_U(\Delta,\delta)$ is of the form $\delta \,\cup\,
% F_U(\Delta)$. 
The transformer ($\T_I(\Delta,\delta)$) for the subsequent \C{insert}
statement can be similarly constructed:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \{\C{ol\_o\_id}=\C{k2};\, 
        \C{ol\_d\_id}=\C{k3};\, \C{ol\_i\_id}=\C{item\_req.ol\_i\_id};\, 
        \C{ol\_qty}=\C{item\_req.ol\_qty}\}
\end{array}
\end{smathpar}
Observe that both transformers are of the form $T(\Delta,\delta) =
\stl \cup F(\stg)$, where $F$ is a function that returns the set of
records added by the transformer to the transaction-local database
($\stl$). Let $\F_U$ and $\F_I$ be the corresponding functions for
$\T_U$ and $\T_I$ shown above. The state transformed of the loop body
in Fig.~\ref{fig:new_order_code} can be expressed as the following
composition of $\F_U$ and $\F_I$:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \F_U(\stg) \cup \F_I(\stg)
\end{array}
\end{smathpar}
The transformer for the loop itself can now be computed to be:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup \C{item\_reqs}\bind
      (\lambda\C{item\_req}.~ \F_U(\stg) \cup \F_I(\stg))
\end{array}
\end{smathpar}
Observe that the structure of the transformer mirrors the structure of
the program itself. In particular, SQL statements become set
operations, and the \C{DB} monad's $\bind$ becoms set monadic $\bind$,
leading to a completely mechanical inference procedure.  The advantage
of inferring such transformers is that we can now make use of a
semantics-preserving translation from the domain of sets equipped with
$\bind$ to first-order logic, allowing us to leverage SMT solvers for
automated proofs without having to infer potentially complex
thread-local invariants or intermediate assertions.
Sec.~\ref{sec:inference} describes this translation. In the exposition
thus far, we assumed $\Delta$ remains invariant, which is clearly not
the case when we admit concurrency.  Necessary concurrency extensions
of the state transformer semantics to deal with interference is also
covered in Sec.~\ref{sec:inference}.  The following two sections,
however, first focus on laying the theoretical foundations for
reasoning about weak isolation.

### 2017-07-07 opsem.tex
Edits.
edits

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). Each transaction is assumed to have a unique
identifier $i$. The $\I$ in the \C{TXN} block syntax is the
transaction's isolation specification, which will be explained later.
Certain terms that only appear at run-time are also present in $c$.
These include a {\sf txn} block tagged with sets ($\stl$ and $\stg$) of
records whose meaning is explained below, and a runtime {\sf foreach}
expression that keeps track of the set ($s_1$) of items already
iterated, and the set ($s_2$) of items yet to be iterated. Note that
even the surface-level syntax of the \C{FOREACH} command shown here is
a little different from the one used in previous sections; its
higher-order function has two arguments, $y$ and $z$, which are
invoked (during the reduction) with the set of already-iterated items,
and the current item, respectively. This form of \C{FOREACH} lends
itself to inductive reasoning that will be useful for verification
(Sec.~\ref{sec:reasoning}).

@@

We define a small-step operational semantics for this language in
terms of an abstract machine that executes a command, and updates
either a transaction-local ($\stl$), or global ($\stg$) database, both
of which are modeled as a set of records of a pre-defined type, i.e.,
they all belong to a single table.  The generalization to multiple
tables is straightforward. The semantics assumes that the records in
$\stg$ are uniquely identifiable through their $\idf$ field, and
enforces this invariant wherever necessary. Certain hidden fields are
treated specially by the operational semantics, and they are hidden
from the surface language. These include a $\txnf$ field that tracks
the identifier of the transaction that last updated the record, and a
$\delf$ field that flags the deleted records in $\stl$.  For a set $S$
of records, we define $\dom(S)$ as the set of unique ids of all
records in $S$. Thus $|\dom(\stg)| = |\stg|$. During its execution, a
transaction may write to multiple records in $\stg$. Atomicity
dictates that such writes should not be visible in $\stg$ until the
transaction commits. We therefore associate each transaction with a
local database ($\stl$) that stores such uncommitted
records\footnote{While SQL's \C{UPDATE} admits writes at the
  granularity of record fields, most popular databases enforce
  record-level locking, allowing us to think of ``uncommitted writes''
  as ``uncommitted records''. }. Uncommitted records include deleted
records, whose $\delf$ field is set to \C{true}. When the transaction
commits, its local database is atomically \emph{flushed} to the global
database, committing these heretofore uncommitted records. The flush
operation ($\gg$) is defined as follows:
\begin{smathpar}
\begin{array}{c}
\forall r.~ r \in (\stl\gg\stg) ~\Leftrightarrow~ 
  (r.\idf \notin \dom(\stl) \conj r \in \stg)
\disj (r \in \stl \conj \neg r.\delf) 
\end{array}
\end{smathpar}
Let $\stg' = \stl\gg\stg$. A record $r$ belongs to $\stg'$ iff it
belongs to $\stg$ and has not been updated in $\stl$, i.e., $r.\idf
\notin \dom(\stl)$, or it belongs to $\stl$, i.e., it is either a new
record, or an updated version of an old record, but the update is not
a deletion ($\neg r.\delf$). Thus, flush defines the result of
atomically applying a transaction's local writes to the global
database.  Besides the commit, flush also helps a transaction read its
own writes. Intuitively, the result of a read operation inside a
transaction must be computed on the database resulting from flushing
the current local state ($\stl$) to the global state ($\stg$). The
abstract machine of Fig.~\ref{fig:txnimp}, however, does not let a
transaction read its own writes. This simplifies the semantics,
without losing any generality, since substituting $\stl\gg\stg$ for
$\stg$ at select places in the reduction rules effectively allows
reads of uncommitted transaction writes to be realized.

@@

\textbf{Local Context Independence} As mentioned previously, our
operational semantics does not let a transaction read its own writes.
It also does not let a transaction overwrite its own writes, due to
the premise $\dom(\stl)\cap\dom(s) = \emptyset$ on
\rulelabel{E-Delete} and \rulelabel{E-Update} rules. We refer to this
restriction as \emph{local context independence}. As such, this
restriction is easy to relax in operational semantics and the
reasoning framework presented in the next section, however, the
automation framework introduced in Sec.~\ref{sec:inference} has a
non-trivial dependence on this assumption. Nonetheless, our experience
suggests that local context independence is not a severe restriction
in practice. For instance, all 17 transactions in the TPC-C and TPC-E
benchmarks satisfy this assumption.

>>>>

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). Each transaction is assumed to have a unique
identifier $i$, and executes at the top-level; our semantics does not
supported nested transactions. The $\I$ in the \C{TXN} block syntax is
the transaction's isolation specification, which will be explained
later.  Certain terms that only appear at run-time are also present in
$c$.  These include a {\sf txn} block tagged with sets ($\stl$ and
$\stg$) of records whose meaning is explained below, and a runtime
{\sf foreach} expression that keeps track of the set ($s_1$) of items
already iterated, and the set ($s_2$) of items yet to be
iterated. Note that even the surface-level syntax of the \C{FOREACH}
command shown here is a little different from the one used in previous
sections; its higher-order function has two arguments, $y$ and $z$,
which are invoked (during the reduction) with the set of
already-iterated items, and the current item, respectively. This form
of \C{FOREACH} lends itself to inductive reasoning that will be useful
for verification (Sec.~\ref{sec:reasoning}).

@@

We define a small-step operational semantics for this language in
terms of an abstract machine that executes a command, and updates
either a transaction-local ($\stl$), or global ($\stg$) database, both
of which are modeled as a set of records of a pre-defined type, i.e.,
they all belong to a single table.  The generalization to multiple
tables is straightforward. The semantics assumes that the records in
$\stg$ are uniquely identifiable through their $\idf$ field, and
enforces this invariant wherever necessary. Certain hidden fields are
treated specially by the operational semantics, and they are hidden
from the surface language. These include a $\txnf$ field that tracks
the identifier of the transaction that last updated the record, and a
$\delf$ field that flags deleted records in $\stl$.  For a set $S$ of
records, we define $\dom(S)$ as the set of unique ids of all records
in $S$. Thus $|\dom(\stg)| = |\stg|$. During its execution, a
transaction may write to multiple records in $\stg$. Atomicity
dictates that such writes should not be visible in $\stg$ until the
transaction commits. We therefore associate each transaction with a
local database ($\stl$) that stores such uncommitted
records\footnote{While SQL's \C{UPDATE} admits writes at the
  granularity of record fields, most popular databases enforce
  record-level locking, allowing us to think of ``uncommitted writes''
  as ``uncommitted records''. }. Uncommitted records include deleted
records, whose $\delf$ field is set to \C{true}. When the transaction
commits, its local database is atomically \emph{flushed} to the global
database, committing these heretofore uncommitted records. The flush
operation ($\gg$) is defined as follows:
\begin{smathpar}
\begin{array}{c}
\forall r.~ r \in (\stl\gg\stg) ~\Leftrightarrow~ 
  (r.\idf \notin \dom(\stl) \conj r \in \stg)
\disj (r \in \stl \conj \neg r.\delf) 
\end{array}
\end{smathpar}
Let $\stg' = \stl\gg\stg$. A record $r$ belongs to $\stg'$ iff it
belongs to $\stg$ and has not been updated in $\stl$, i.e., $r.\idf
\notin \dom(\stl)$, or it belongs to $\stl$, i.e., it is either a new
record, or an updated version of an old record, but the update is not
a deletion ($\neg r.\delf$). Thus, flush defines the result of
atomically applying a transaction's local writes to the global
database.  Besides the commit, flush also helps a transaction read its
own writes. Intuitively, the result of a read operation inside a
transaction must be computed on the database resulting from flushing
the current local state ($\stl$) to the global state ($\stg$). The
abstract machine of Fig.~\ref{fig:txnimp}, however, does not let a
transaction read its own writes. This simplifies the semantics,
without losing any generality, since substituting $\stl\gg\stg$ for
$\stg$ at select places in the reduction rules effectively allows
reads of uncommitted transaction writes to be realized.

@@

\textbf{Local Context Independence} As mentioned previously, our
operational semantics does not let a transaction read its own writes.
It also does not let a transaction overwrite its own writes, due to
the premise $\dom(\stl)\cap\dom(s) = \emptyset$ on
\rulelabel{E-Delete} and \rulelabel{E-Update} rules. We refer to this
restriction as \emph{local context independence}. As such, this
restriction is easy to relax in operational semantics and the
reasoning framework presented in the next section, however, the
automation framework introduced in Sec.~\ref{sec:inference} has a
non-trivial dependence on this assumption. Nonetheless, we haven't
found local context independence to be a severe restriction in
practice. For instance, all 17 transactions in the TPC-C and TPC-E
benchmarks satisfy this assumption.
